---
title: "PreProcessing"
author: "Yevashan Perumal"
date: '2022-06-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Imports
```{r}
# library imports
library(tidyverse)
library(gridExtra)
library(Amelia)
```

## Data Import
```{r}
# Mac
# df <- read.csv('/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/data with some cdw.csv')

# Windows
df <- read.csv('C:/Users/X464371/OneDrive - University of Cape Town/Yevashan Perumal MSc - OneDrive/Data/data with some cdw.csv')
```


## Columns dropped from EDA

Setup Variables/Lists used below
```{r}
# dropped columns from EDA
cols_to_drop <- c(# Just didnt seem handy:
                  'RecordID',
                  'ProductCode',
                  'DimGCSPartyKeyLifeCovered',
                  'DimDateKeyInitiated','DimDateKeyFinalised','DimDateKeyAccepted'
                  ,'UnderwritingInitiationDate','UnderwritingFinaliseDate',
                  'UnderwritingCode',
                  'UnderwritingTerritory',
                  'UnderwritingDuration',
                  'QuoteNumber' ,
                  # because of high correlation/target leakage/they're informed by GammaGT:
                  'LoadingPercentage','CoverLoadingAmount','HealthLoadingPercentage',
                  # double check these are all for current app?could be previous:
                  'CoverLoading','HealthLoadingIndicator','CoverLoadingIndicator',
                  'UnderwritingOutcome','NoOfLoadingsApplied',
                  #customer identifier:
                   'LifeCoveredClientID',
                  # no varation, found in modelling stage
                  'NumberOfVerifications'
                  )
```
The columns above just looked useless from eyeballing and seeing what they are. Dropping duplicates because there were multiple things again.

## Converting char to numeric for appropriate columns
```{r}
char_to_numeric_list <- c( 
'CurrentCoverAmount','CholesterolValue','GammaGTValue','CotinineValue',
'BloodSugarValue','ClientBMIValue','DoctorBMIValue','SystolicBloodPressureValue',
'DiastolicBloodPressureValue','RestingPulseValue','HeightValue','WeightValue',
'NumberOfCigarettes','NumberOfFailedLevel1Evaluations','NumberOfFailedDetailEvaluations','NumberOfExclusions',  'NumberOfDeferrals','MonthlyIncome','MonthlyIncome_Override',
'UnderwritingCreditAmount','InputMedium','CoverAmount',
'PremiumRateCatDiscount','ReinsuranceRateCatDiscount','IntermediaryQualityDiscount',  
'MonthlyIncome_Override2','AGE','ALL_BEN_COUNT','ALL_CONTRACT_COUNT','ANNUAL_INVESTMENT','AUM'
)            
```

## Drop some columns after EDA
```{r}
post_eda_drop_cols <- c('CotinineValue','NumberOfCigarettes','NumberOfExclusions',
                        'UnderwritingCreditAmount','InputMedium','ManualDiscountIndicator',
                        'MonthlyIncome_Override2','MonthlyIncome',
                        'PremiumRateCatDiscount',#not a lot of variation here, about 1k rows
                        'ReinsuranceRateCatDiscount', #not a lot of variation here, about 1k rows
                        'IntermediaryQualityDiscount', #not a lot of variation here, about 1k rows
                        'AutoDeclinedIndicator', #cat column, not much variation
                        'AutoUnderwrittenIndicator', #cat column, not much variation
                        'UnderwritingPostalCode', # lots of nulls
                        'IncomeCoverBand', # everything just 100k>
                        'IntermediaryQualityCategory' ,# empty
                        'PUBLIC_SECTOR_SOURCE', # just didnt think it would be handy
                        'ManualDiscountIndicator', #empty
                        'OccupationDescription' #text description
                        )
```

## Columns to convert to factor
```{r}
t1 <-  c("ProductSoldType","SmokerStatus","AutoDeferredIndicator","NTUIndicator","LOAIndicator","ExclusionIndicator",    "AcceleratorIndicator","ASS_ETHNIC_GROUP","PROVINCE","FARMER_IND","PROF_MARKET_IND","PUBLIC_SECTOR_IND",
"PUBLIC_SECTOR_TYPE","MARITAL_STATUS","GENDER")   
```



## Remove Outliers Function
```{r}
# create detect outlier function
detect_outlier <- function(x) {
 
    # calculate first quantile
    Quantile1 <- quantile(x, probs=.25)
 
    # calculate third quantile
    Quantile3 <- quantile(x, probs=.75)
 
    # calculate inter quartile range
    IQR = Quantile3-Quantile1
 
    # return true or false
    x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
}
 
# create remove outlier function
remove_outlier <- function(dataframe,
                            columns=names(dataframe)) {
 
    # for loop to traverse in columns vector
    for (col in columns) {
 
        # remove observation if it satisfies outlier function
        dataframe <- dataframe[!detect_outlier(dataframe[[col]]), ]
    }
 
    # return dataframe
    print("Remove outliers")
    print(dataframe)
}
 

```


```{r}
data_cleaning_long <- function(data){
    # dropped columns from EDA
    data <- data %>% select(-all_of(cols_to_drop)) %>% distinct()
    
    ## Converting char to numeric for appropriate columns
    data <- data %>% mutate_at(char_to_numeric_list, as.numeric)
    
    ## Drop some columns after EDA
    data <- data %>% select(-all_of(post_eda_drop_cols))
    
    ## Deal with nulls generated in the numeric conversion
    data <- data %>% filter(is.na(GammaGTValue)==F)
    data <- data %>% filter(is.na(CurrentCoverAmount)==F)
    data <- data %>% filter(is.na(RestingPulseValue)==F)
    data <- data %>% filter(is.na(NumberOfFailedLevel1Evaluations)==F)
    data <- data %>% filter(is.na(NumberOfDeferrals)==F)
    data <- data %>% filter(is.na(MonthlyIncome_Override)==F)
    
    ## Fill Nulls with appropriate values for numeric and character columns
    data$NumberOfFailedDetailEvaluations[is.na(data$NumberOfFailedDetailEvaluations)]= 0
    data <- data %>% mutate(EducationID_Override = replace(EducationID_Override, EducationID_Override == 'NULL', '1'))
    data <- data %>% mutate(ASS_ETHNIC_GROUP = replace(ASS_ETHNIC_GROUP, ASS_ETHNIC_GROUP == '', 'UNKNOWN'))
    data <- data %>% mutate(PROVINCE = replace(PROVINCE, PROVINCE == '', 'UNKNOWN'))
    data <- data %>% mutate(FARMER_IND = replace(FARMER_IND, FARMER_IND == '', 'UNKNOWN'))
    data <- data %>% mutate(PROF_MARKET_IND = replace(PROF_MARKET_IND, PROF_MARKET_IND == '', 'UNKNOWN'))
    data <- data %>% mutate(PUBLIC_SECTOR_IND = replace(PUBLIC_SECTOR_IND, PUBLIC_SECTOR_IND == '', 'UNKNOWN'))
    data <- data %>% mutate(PUBLIC_SECTOR_TYPE = replace(PUBLIC_SECTOR_TYPE, PUBLIC_SECTOR_TYPE == '', 'UNKNOWN'))
    data <- data %>% mutate(PUBLIC_SECTOR_TYPE = replace(PUBLIC_SECTOR_TYPE, PUBLIC_SECTOR_TYPE == 'N/A', 'UNKNOWN'))
    data <- data %>% mutate(MARITAL_STATUS = replace(MARITAL_STATUS, MARITAL_STATUS == '', 'UNKNOWN'))
    data <- data %>% mutate(GENDER = replace(GENDER, GENDER == '', 'UNKNOWN'))
    
    # Find all numeric columns 
    num_cols <- data %>% select_if(is.numeric)
    # Find all char columns
    char_cols <- data %>% select_if(is.character)
    
    # Outliers
    data <- remove_outlier(data, columns = names(num_cols %>%
                                            select(-AGE,-ALL_BEN_COUNT,-ALL_CONTRACT_COUNT,-ANNUAL_INVESTMENT,-AUM)))
    
    #remove duplicate rows
    data <- data %>% distinct()
    
    # coerce char columns to factor, no order
    data[t1] <- lapply(data[t1], factor) 
    
    # "OccupationClassCode"  
    data$OccupationClassCode <- factor(data$OccupationClassCode, order = TRUE, 
                                        levels = c("-","A", "B", "C","D","E"))
    
    # "DisabilityIncomeClass"  
    data$DisabilityIncomeClass <- factor(data$DisabilityIncomeClass, order = TRUE, 
                                        levels = c("Unknown","1", "1*", "2","3"))
    
    # "EducationID_Override"  
    data$EducationID_Override <- factor(data$EducationID_Override, order = TRUE, 
                                        levels = c("-1", "1", "2","3","4"))
    
     # "SumAssuredBand"
    data$SumAssuredBand <- factor(data$SumAssuredBand, order = TRUE, 
                                        levels = c( "A: R0.00 - R100,000.00"   , 
                                                    "B: R100,000.00 - R200,000.00"  , 
                                                    "C: R200,000.00 - R300,000.00"    ,
                                                     "D: R300,000.00 - R400,000.00"  ,
                                                     "E: R400,000.00 - R500,000.00"   ,
                                                     "F: R500,000.00 - R600,000.00" ,
                                                     "G: R600,000.00 - R700,000.00" ,
                                                    "H: R700,000.00 - R800,000.00"  ,
                                                     "I: R800,000.00 - R900,000.00"  ,
                                                       "J: R900,000.00 - R1,000,000.00"  ,
                                                      "K: R1,000,000.00 - R1,500,000.00",
                                                     "L: R1,500,000.00 - R2,000,000.00",
                                                     "M: R2,000,000.00 - R2,500,000.00"
                                                    ))

    
    return(data) 

}
```

```{r}
t <- data_cleaning_long(df)
```

Deal with highly correlated variables(should've done a corrplot in the EDA)
```{r}
# library(caret)
# df_corr = cor(data %>% select_if(is.numeric) %>% select(-AGE,-ALL_BEN_COUNT,-ALL_CONTRACT_COUNT,-ANNUAL_INVESTMENT,-AUM))
# # df_corr = cor(data)
# 
# hc = findCorrelation(df_corr, cutoff=0.6) # putt any value as a "cutoff"
# hc = sort(hc)
# reduced_Data = df1[,-c(hc)]
# print (reduced_Data)
```


At this points only the CDW numeric columns have null values remaining.
Going to be making two datasets now:

```{r}
# dataset with CDW columns, but now nulls
df_wide <- data %>% na.omit()
dim(df_wide)

write.csv(df_wide,"/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/wide_data.csv", row.names=FALSE)

```

```{r}
# dataset dropping CDW columns, but with more rows. 
df_long <- data %>% select(-AGE,-ALL_BEN_COUNT,-ALL_CONTRACT_COUNT,-ANNUAL_INVESTMENT,-AUM)

dim(df_long)

write.csv(df_long,"/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/long_data.csv", row.names=FALSE)
```

```{r}
save(df_wide,df_long,file = "/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/data_for_modelling.Rdata")
```

