
To Do:

remake the wide and long from the imputed set by using indices.Actually will only be usefull for the long dataset

decided to
wide: complete case
long: imputing data with a few missing cases
all: missing data where some columns have a lot of missingness

long and all will have same number of rows

8000 rows test



Research objective:
do the different types of data processing methods yielded by imputatuion affect model accuracy differently for the different models?


# Import Libraries
```{r message=FALSE, warning=FALSE}
library(xgboost)
library(tidyverse)
library(caret)
library(randomForest)
library(ggplot2)
library(rpart)
library(tree)

#Imputation
library(mice)
library(Amelia)
library(Rcpp)
library('fastDummies')
library(missForest)
```

# Import Data

```{r}
# Mac
# load("/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/data_for_modelling.Rdata")

# Windows load tables saved as Rdata to preserve types
load("C:/Users/X464371/OneDrive - University of Cape Town/Yevashan Perumal MSc - OneDrive/Data/data_for_modelling.Rdata")

# load images
# load("/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/imput_splits_image.Rdata")
```

# Some more cleaning
```{r}
#finding zero variance columns; was an issue during imputation

# zv <- apply(df_long %>% select_if(is.numeric), 2, function(x) length(unique(x)) == 1)
# sum(zv)
# zv

# Find names of highly correlated columns
# df_corr = cor(df_long %>% select_if(is.numeric))
# hc = findCorrelation(df_corr, cutoff=0.6) # putt any value as a "cutoff"
# hc = sort(hc)
# #the name of the columns chosen above
# to_be_removed <- colnames(df_corr)[hc]
```


```{r}
# Remove highly correlated columns
# Not sure if all should be removed or just some.

# df_long <- df_long %>% select(-CurrentCoverAmount,-WeightValue,-SystolicBloodPressureValue,-NumberOfDeferrals,-ClientBMIValue)

df_wide <- df_wide %>% select(-CurrentCoverAmount,-WeightValue,-SystolicBloodPressureValue,-NumberOfDeferrals,-ClientBMIValue) # might need to do the same exercise as above when I get to the imputation

to_be_imputed <- to_be_imputed %>% select(-CurrentCoverAmount,-WeightValue,-SystolicBloodPressureValue,-NumberOfDeferrals,-ClientBMIValue)

# drop the cols with lots of missing values; the new 'long' dataset
 to_be_imputed_less_cols <- to_be_imputed %>%select(-AGE,-ALL_BEN_COUNT,-ALL_CONTRACT_COUNT,
                                                                   -ANNUAL_INVESTMENT,-AUM)


# dim(df_long)
# dim(df_wide)
# dim(to_be_imputed)
```

```{r}
# Visualise missingness
md.pattern(to_be_imputed)
```



# Imputation 

 
## Random MissForest
```{r}
# all data imputatioan
miss_forest_imp <- missForest(to_be_imputed,maxiter=5,ntree=5)
miss_forest_imputed_all_df <-  miss_forest_imp$ximp
```

```{r}
# long dataset imputation
miss_forest_imp_less_cols <- missForest(to_be_imputed_less_cols,maxiter=5,ntree=5)
miss_forest_imputed_less_cols_df <-  miss_forest_imp_less_cols$ximp
```


```{r}
sum(is.na(miss_forest$ximp))
```

## MICE
```{r}
# system.time(
# imp <- mice(imput_data,seed = 42) 
# )
imp <- mice(to_be_imputed,method='cart',seed = 42,m=5,maxit=5)
# imp <- mice(head(to_be_imputed,1000),seed = 42) 


sapply(imp$data, function(x) sum(is.na(x)))
# this works
# test <- imput_data %>% select_if(is.numeric)
# imp <- mice(test,seed = 42)

# imp$imp$GammaGTValue
# meth = init$method
# predM = init$predictorMatrix
```
Error here. looks to be with highly correlted variables due to the factor vars I have. Need to either decide to remove, or use different methods for those ones.
looks like it starts acting up at more than 31500 rows. Need to investigate further what's gonig on here.
 
 
```{r}
# first imputed dataset
complete(imp,1)
```
```{r}
# second imputed dataset
complete(imp,2)
```
 
```{r}
plot(imp)
```
 
```{r}
# diagnostic checking to see if imputed values match
stripplot(imp, chl~.imp, pch=20, cex=2)
```
 

```{r}
summary(to_be_imputed)
```
 
```{r}
summary(complete(imp))
```
compare means in the two summaries
 
 
```{r}
pool(imp)
```


## Amelia
assumes varaibles follow multivariate normal distributions. Some transfaormations may be needed?
```{r}
imput_data_dummy_var <- dummy_cols(to_be_imputed, select_columns = c('ProductSoldType', 'SmokerStatus', 'OccupationClassCode', 'DisabilityIncomeClass', 'AutoDeferredIndicator', 'NTUIndicator', 'LOAIndicator', 'ExclusionIndicator', 'AcceleratorIndicator', 'EducationID_Override', 'SumAssuredBand', 'ASS_ETHNIC_GROUP', 'PROVINCE', 'FARMER_IND', 'PROF_MARKET_IND', 'PUBLIC_SECTOR_IND', 'PUBLIC_SECTOR_TYPE', 'MARITAL_STATUS', 'GENDER'),
           remove_selected_columns = TRUE)
```
 
```{r}
set.seed(42)
amelia(imput_data_dummy_var %>% select(-'ProductSoldType_Standalone With Accelerator', -SmokerStatus_SMOKER, -OccupationClassCode_E, -DisabilityIncomeClass_3, -AutoDeferredIndicator_Y, -NTUIndicator_Y, -LOAIndicator_Y, -ExclusionIndicator_Y,- AcceleratorIndicator_Y, -EducationID_Override_4, -`SumAssuredBand_M: R2,000,000.00 - R2,500,000.00`, -ASS_ETHNIC_GROUP_WHITE, -`PROVINCE_W CAPE`,- FARMER_IND_Y, -PROF_MARKET_IND_UNKNOWN, -PROF_MARKET_IND_Y, -PUBLIC_SECTOR_IND_UNKNOWN, -PUBLIC_SECTOR_IND_Y, -`PUBLIC_SECTOR_TYPE_WESTERN-CAPE PROVINCIAL GOVERMENT`, -MARITAL_STATUS_WIDOWED, -GENDER_MALE,-GENDER_UNKNOWN),m=5)


amelia(to_be_imputed,m=1)
```
##########################################################################################################################

# Train-Test Split
```{r}
# Create the training and test datasets
set.seed(42)

# Step 1: Get row numbers for the training data
trainRowNumbers <- createDataPartition(df_wide$GammaGTValue, p=0.74, list=FALSE)

# Step 2: Create the training  dataset
wide_train_df <- df_wide[trainRowNumbers,]

wide_x_train <- wide_train_df %>% select(-GammaGTValue)
wide_y_train <- wide_train_df %>% select(GammaGTValue)

# Step 3: Create the test dataset
wide_test_df <- df_wide[-trainRowNumbers,]

wide_x_test <- test_df %>% select(-GammaGTValue)
wide_y_test <- test_df %>% select(GammaGTValue)
```

```{r}
# This is the test DF for all the models to be tested on
mf_test_df <- miss_forest_imputed_all_df[tibble::rownames_to_column(wide_test_df)$rowname,]

mf_train_df <-  anti_join(tibble::rownames_to_column(miss_forest_imputed_all_df),tibble::rownames_to_column(mf_test_df),by=c("rowname" = "rowname"))

```

```{r}
mf_train_df <- miss_forest_imputed_all_df[trainRowNumbers,]
mf_test_df <- miss_forest_imputed_all_df[-trainRowNumbers,]
```
```



```{r}
save.image("/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/imput_splits_image.Rdata")
```