---
title: 'R Notebook'
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---
## Import packages
```{r message=FALSE, warning=FALSE}
# library imports
library(tidyverse)
library(gridExtra)
library(Amelia)
library(corrplot)   
```

## Import Data

```{r}
# df <- read.csv('/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/data with some cdw.csv')

# df <- read.csv('/Users/yevashanperumal/Library/CloudStorage/OneDrive-UniversityofCapeTown/Yevashan Perumal MSc - OneDrive/Data/data with some cdw.csv')

df <- read.csv('D:/Users/X464371/OneDrive - University of Cape Town/Yevashan Perumal MSc - OneDrive/Data/data with some cdw.csv')
```

```{r}
head(df)
```

```{r}
dim(df)
```

Drop the unnecessary columns

```{r}
cols_to_drop <- c(# Not useful upon inspections:
                  'RecordID',
                  'ProductCode',
                  'DimGCSPartyKeyLifeCovered',
                  'DimDateKeyInitiated','DimDateKeyFinalised','DimDateKeyAccepted'
                  ,'UnderwritingInitiationDate','UnderwritingFinaliseDate',
                  'UnderwritingCode',
                  'UnderwritingTerritory',
                  'UnderwritingDuration',
                  'QuoteNumber' ,
                  # because of high correlation/target leakage/they're informed by GammaGT:
                  'LoadingPercentage','CoverLoadingAmount','HealthLoadingPercentage',
                  'CoverLoading','HealthLoadingIndicator','CoverLoadingIndicator',
                  'UnderwritingOutcome','NoOfLoadingsApplied',
                  #customer identifier:
                   'LifeCoveredClientID'  
                  )

#drop columns
df2 <- df %>% select(-all_of(cols_to_drop))

#remove duplicate rows
df2 <- df2 %>% distinct()

dim(df2)
```

```{r}
str(df2)
```

Lots of numeric values coming through as characters; More nulls will
likely be introduced when I convert them

```{r}
char_to_numeric_list <- c( 
'CurrentCoverAmount','CholesterolValue','GammaGTValue','CotinineValue',
'BloodSugarValue','ClientBMIValue','DoctorBMIValue','SystolicBloodPressureValue',
'DiastolicBloodPressureValue','RestingPulseValue','HeightValue','WeightValue',
'NumberOfCigarettes','NumberOfFailedLevel1Evaluations','NumberOfFailedDetailEvaluations','NumberOfExclusions',    
'NumberOfVerifications','NumberOfDeferrals','MonthlyIncome','MonthlyIncome_Override',
'UnderwritingCreditAmount','InputMedium','CoverAmount',
'PremiumRateCatDiscount','ReinsuranceRateCatDiscount','IntermediaryQualityDiscount',  
'MonthlyIncome_Override2','AGE','ALL_BEN_COUNT','ALL_CONTRACT_COUNT','ANNUAL_INVESTMENT','AUM'
)            
df2 <- df2 %>% mutate_at(char_to_numeric_list, as.numeric)
```

```{r}
missmap(df2)
```

```{r}
colSums(is.na(df2))
```

-Couple columns I could definitely drop with inspection: - ManualDiscountIndicator - NumberOfVerifications -
CotinineValue - InputMedium

-Some missing values for GammaGT; 

```{r}
# Unique values for all variables in the dataframe

# tab1 <- df2 %>% summarise_all(n_distinct)%>%t()
# tab1 <- data.frame(tab1)
# tab1 %>% arrange(tab1)
```

# Univariate exploration

## Numeric Variables (and ones that should be)

```{r}
box_and_hist <- function(variable){
    hist <- df2%>%ggplot(aes_string(x=variable))+geom_histogram()+
            theme(axis.title.x=element_blank(),
                  axis.title.y = element_text(size = 6),
                  plot.title = element_text(size = 8))+
            labs(title = paste('Histogram and Boxplot of',variable))
    boxp <- df2%>%ggplot()+
      geom_boxplot(aes_string(x=variable))+
      theme(axis.title.x = element_text(size =8))
    grid.arrange(hist,boxp,nrow=2)
    
    summary(df2 %>% select(variable))
}

# add option to create a filtered view?
```

```{r}
countplot_explore <- function(input_variable){
    bar <- df2%>%ggplot(aes_string(x=input_variable))+geom_bar()+
      labs(title = paste('Countplot of ',input_variable))
    print(input_variable)
    print(sum(is.na(df2[input_variable])))
    bar
    # summary(df2 %>% select(variable))
}

# add option to create a filtered view?
```

```{r message=FALSE, warning=FALSE}
num_cols <- df2 %>% select_if(is.numeric) %>% colnames()

char_cols <- df2 %>% select_if(is.character) %>% colnames()
```

------------------------------------------------------------------------

```{r}
# CurrentCoverAmount
box_and_hist(num_cols[1])
```

```{r}
# CholesterolValue
box_and_hist(num_cols[2])
```

Couple large outliers that need to be cleaned(winsorised an option?).
Probably data capture errors.


```{r}
# CotinineValue
box_and_hist(num_cols[4])
```

Dont think I can use this one by the number of missing values. I remember being advised that I should use the SmokerStatus ind instead.

```{r}
# BloodSugarValue
box_and_hist(num_cols[5])
```

Need to take care of very large outliers. And the zero values.

```{r}
# ClientBMIValue
box_and_hist(num_cols[6])
```

Need to take care of very large outliers. And the very small(impossible)
values.

```{r}
# DoctorBMIValue
box_and_hist(num_cols[7])
```

Same as above

```{r}
# DoctorBMIValue
box_and_hist(num_cols[8])
```

Large outliers. Seems like there may be a lot of zero values; not sure
if it affects the quality of the feature.

```{r}
# DoctorBMIValue
box_and_hist(num_cols[9])
```

Same as systolic by the looks of it

```{r}
# RestingPulseValue
box_and_hist(num_cols[10])
```

Outliers.

```{r}
# HeightValue
box_and_hist(num_cols[11])
```

Outliers again. Maybe people using different scales. Probably a high
correlation with BMI; not as much of an issue for tree based methods.
Might need to be removed anyway?

```{r}
# WeightValue
box_and_hist(num_cols[12])
```

Same as height

```{r}
# NumberOfCigarettes
box_and_hist(num_cols[13])
```

Useless. All zero, need to drop.

```{r}
# NumberOfFailedLevel1Evaluations
box_and_hist(num_cols[14])
```

Not terrible. Need to confirm what this feature means.

```{r}
# NumberOfFailedLevel1Evaluations
box_and_hist(num_cols[15])

#countplot
```

Same as above. Missing for about half the data though.

```{r}
# NumberOfExclusions
box_and_hist(num_cols[16])

# dim(df[which(df2$NumberOfExclusions>0),]) # only 5987 greater than zero
```

This one also seems like it might be full of zeroes and need to be
dropped.

```{r}
# NumberOfVerifications
box_and_hist(num_cols[17])

#countplot?
```

Needs to confirm what this one means

```{r}
# NumberOfDeferrals
box_and_hist(num_cols[18])
```

Also need to check what this means.

```{r}
# MonthlyIncome
box_and_hist(num_cols[19])
```

Outliers skewing it. And about 16k missing values.

```{r}
# MonthlyIncome
box_and_hist(num_cols[20])
```

Max value here looks like nonsense. Less missing values than income fields above though.

```{r}
# UnderwritingCreditAmount
box_and_hist(num_cols[21])
```

All zeroes; needs to drop

```{r}
# IRPCode
box_and_hist(num_cols[22])

# Countplot?This looks like a category
```

```{r}
# InputMedium
box_and_hist(num_cols[23])
```

All nulls, need to drop.

```{r}
# EntryANB
box_and_hist(num_cols[24])
```

Seems usable, but now to confirm definition

```{r}
# CoverAmount
box_and_hist(num_cols[25])
```

Big values, but those are possible. Maybe banding this to take care of
them e.g. \>R5mil (maybe consider for above values too)

```{r}
# EducationID
box_and_hist(num_cols[26])

# This looks like a category fields too
```

```{r}
# PremiumRateCatDiscount
box_and_hist(num_cols[27])
```

Lots of nulls.lots of zeroes. also not sure if it is something
determined by gamma GT/med tests. But think it might be a drop. Looks
like they go in 2.5% jumps. Category field like Wang2021 did?

```{r}
# ReinsuranceRateCatDiscount
box_and_hist(num_cols[28])
```

Same as above. Lots of zeroes and nulls.

```{r}
# IntermediaryQualityDiscount
box_and_hist(num_cols[29])
```

Seems more useful than above two, but still lots of zero vallues. Maybe
an indicator column and then see how many non zeroes?

```{r}
# ManualDiscountIndicator
box_and_hist(num_cols[30])

 # Actually check if the field is a Y/N?
unique(df$ManualDiscountIndicator)
```

Empty, drop. Actually check if the field is a Y/N?

```{r}
# MonthlyIncome_Override2
box_and_hist(num_cols[31])
```

MonthlyIncome_Override2. need to check if this is a duplicate of
MonthlyIncome_Override; maybe came from different tables.

```{r}
# AGE
box_and_hist(num_cols[32])
```

Key variable. OUtliers again.

```{r}
# ALL_BEN_COUNT
box_and_hist(num_cols[33])
```

CDW; lots of nulls though

```{r}
# ALL_CONTRACT_COUNT
box_and_hist(num_cols[34])
```

Same as above

```{r}
# ANNUAL_INVESTMENT
box_and_hist(num_cols[35])
```

Same as above;

```{r}
# AUM
box_and_hist(num_cols[36])
```

Same as above

------------------------------------------------------------------------

## Categorical Plots

```{r}
# ProductSoldType
countplot_explore(char_cols[1])
```

Can probably group Acell Only Not sure it would be super relevant to
Gamma GT

```{r}
# SmokerStatus
countplot_explore(char_cols[2])
```

Def use.

```{r}
# ProductSoldType
countplot_explore(char_cols[3])

sum(is.na(df2$OccupationClassCode))
```

Def use. Important from underwriting criteria perspective

```{r}
# OccupationDescription
# char_cols[4]
head(df$OccupationDescription)
sum(is.na(df$OccupationDescription))
```

Lots of text here. Maybe could feature engineer with some nlp? could be interesting;

```{r}
# DisabilityIncomeClass
countplot_explore(char_cols[5])
```

Could be an interesting var. Some spread of data, but majority unknown.

```{r}
# AutoDeclinedIndicator
countplot_explore(char_cols[6])
```

Doesnt look  helpful.Drop

```{r}
# AutoUnderwrittenIndicator
countplot_explore(char_cols[7])
```

Same as above.

```{r}
# ProductSoldType
countplot_explore(char_cols[8])
```

This one looks like it could be more helpful. Think I can use it.

```{r}
# ProductSoldType
countplot_explore(char_cols[9])
```

Customer had previous NTU more likely to have higher/lower gamma GT?
Need to double check the definition of this field.

```{r}
# ProductSoldType
countplot_explore(char_cols[10])
```

Seems useful, but need to check what this is.

```{r}
# ProductSoldType
# countplot_explore(char_cols[11])

head(df$UnderwritingPostalCode)
length(unique(df$UnderwritingPostalCode))
```

Proxy for location perhaps. could be handy. I do need to double check how
many were recorded as 'Unknown' to see if this is very usable.

```{r}
# ExclusionIndicator
countplot_explore(char_cols[12])
```

```{r}
# ProductSoldType
countplot_explore(char_cols[13])
```

Looks better populated than that other product field I had above.
Maybe I can drop that and keep this.

```{r}
# EducationID_Override
countplot_explore(char_cols[14])
```

Check with the normal education fields which to use. Maybe I can just
combine them, and when they dont match take this one.

```{r}
# ProductSoldType
countplot_explore(char_cols[15])

unique(df$IncomeCoverBand)
```

Coverage here looks weird. Need to compare with the other income fields
and see what I should use for Income.

```{r}
# ProductSoldType
countplot_explore(char_cols[16])
```

```{r}
# ProductSoldType
countplot_explore(char_cols[17])
```

Empty, drop.

```{r}
# ProductSoldType
countplot_explore(char_cols[18])
```

From CDW; looking good.

```{r}
# PROVINCE
countplot_explore(char_cols[19])
```

From CDW; looking good.

```{r}
# FARMER_IND
countplot_explore(char_cols[20])
```

Some farmers up in here yo. Unknowns up in here too.

```{r}
# PROF_MARKET_IND
countplot_explore(char_cols[21])
```

Also some professionals around town.Unknowns too.

```{r}
# PUBLIC_SECTOR_IND
countplot_explore(char_cols[22])
```

Also some civil servants around town.Unknowns too.

```{r}
# ProductSoldType
countplot_explore(char_cols[23])
```

Some values. Need to group some of these together for sure.

```{r}
# ProductSoldType
countplot_explore(char_cols[24])
```

Fair chunk of NA in here, but think could be usable. NOt sure how it's
handy related to GammaGT though.

```{r}
# ProductSoldType
countplot_explore(char_cols[25])
```

Marriage=high blood pressure? OK,boomer.

```{r}
# ProductSoldType
countplot_explore(char_cols[26])
```

More unknowns than expected.

```{r}
# 
countplot_explore('ManualDiscountIndicator')
```
Eh seems pretty empty; probably drop.

# Target Variable

```{r}
df2$GammaGTValue <-  as.integer(df2$GammaGTValue)
df2 %>% filter(GammaGTValue<200) %>%  ggplot(aes(x=GammaGTValue))+geom_histogram(binwidth = 10)
```

Some really weird large outliers than need to be taken care of. Also
some NAs introduced when converting to character.

# End

# Plots for the write up
```{r}
(raw_hist <- df2 %>% ggplot(aes(x=GammaGTValue))+
    geom_histogram(binwidth = 10)+
    theme_minimal()+
    theme(aspect.ratio=1)+
    # ggtitle('Histogram of Unprocessed GGT Distribution')+
    xlab("GGT Value (IU/L)")+
    ylab("Count")
)


ggsave('raw_hist.pdf',device = "pdf", raw_hist,path ="D:\\Users\\X464371\\OneDrive - University of Cape Town\\Yevashan Perumal MSc - OneDrive\\Write-up\\Figures",width = 5, height = 5)
```

```{r}
df2 %>% ggplot(aes(x=GammaGTValue))+geom_boxplot()
```

